---
title: "Calculation of climate trends in the Bavarian Alps"
author: "Oliver Schnuck (oliver.schnuck@br.de), BR Data"
output: 
  html_document
---

This document describes the calculation of climate trends at different wheather stations in the Bavarian Alps, that we published in our project [Schnee Von Morgen](http://schnee-von-morgen.br.de). The results of our calculations are thus reproducible by knitting `./index.Rmd`, which is also part of the [archive](http://schnee-von-morgen.br.de/recherchetool/calculations.zip) including the relevant CSV-files.

## Data description

The files in the input folder serve as input for the calculations described below:

* `./input/klima.csv` contains the average values of temperature, snowheight and snowdays in the winter half-year since 1961. The numbers trace back to the [daily values](ftp://ftp-cdc.dwd.de/pub/CDC/observations_germany/climate/daily/kl/) measured by the DWD.
* `./input/trends.csv` contains the beginning and end of the periods, for which the trend and its significance is calculated. The longest period with no more than 2 missing values per winter half-year since 1961 is chosen.
* `./input/stationIDs.csv` gives information about the wheather stations at which these indicators were measured. 

The files in the output folder contain the results of the process after the calculations are finished. They are identical to the ones, which can be downloaded as zip-file in the info-layer of our [application](http://schnee-von-morgen.br.de/daten).

* `./output/klima.csv` is identical to the input file above.
* `./output/trends.csv` contains the values of the indicators at the beginning / end of each period under consideration - according to the calculated linear regression.
* `./output/stationIDs.csv` ressambles the input file, complemented with the significance values of each indicator according to the Mann-Kendall-Trends.

## Calculations

First of all we load the csv-files that serve as input for the calculations.

```{r}
climate <- read.csv("./input/klima.csv", na.strings=".", stringsAsFactors = FALSE)
stations <- read.csv("./input/stationIDs.csv", stringsAsFactors = FALSE)
trends <- read.csv("./input/trends.csv", stringsAsFactors = FALSE)
```

```{r, echo=FALSE, warning=FALSE, message=FALSE}

knitr::opts_chunk$set(fig.width=12, fig.height=8,  warning=FALSE)

options(scipen = 10)
# load necessary packages
if(!require(dplyr)) {
  install.packages("dplyr", repos="http://cran.us.r-project.org")
  require(dplyr)
}
if(!require(tidyr)) {
  install.packages("tidyr", repos="http://cran.us.r-project.org")
  require(tidyr)
}
if(!require(magrittr)) {
  install.packages("magrittr", repos="http://cran.us.r-project.org")
  require(magrittr)
}
if(!require(jsonlite)) {
  install.packages("jsonlite", repos="http://cran.us.r-project.org")
  require(jsonlite)
}

if(!require(Kendall)) {
  install.packages("Kendall", repos="http://cran.us.r-project.org")
  require(Kendall)
}
```

After having loaded necessary libraries in the background, we calculate the linear regression and the Mann-Kendall-Trend-Tests for each station and the period of maximum length. 

```{r}
perStation <- climate %>% 
  left_join(stations, by = c("id" = "key")) %>%
  filter(date<=trendEnde) %>%
  group_by(id) %>% 
  arrange(date) %>%
  do(
    tempera1 = lm(temperatur ~ seq(1961, 1960 + length(temperatur)), data= .),
    tempera2 = MannKendall(ts(.$temperatur, 1961, 1960 + length(.$temperatur), 1)),
    schneeh1 = lm(schneehoehe ~ seq(1961, 1960 + length(schneehoehe)), data= .),
    schneeh2 = MannKendall(ts(.$schneehoehe, 1961, 1960 + length(.$schneehoehe), 1)),  
    schneet1 = lm(schneetage ~ seq(1961, 1960 + length(schneetage)), data= .),
    schneet2 = MannKendall(ts(.$schneetage, 1961, 1960 + length(.$schneetage), 1))  
  ) 
```

Now we're extracting the slope and the intercept from the linear models and the significance value of the 2-sided Mann-Kendall-Tests.

```{r}
perStation <- perStation %>%  
  mutate(
    slope_tempera=summary(tempera1)$coeff[2],
    intercept_tempera=summary(tempera1)$coeff[1],
    S_tempera= tempera2$S,
    varS_tempera = tempera2$varS,
    p_value_tempera=tempera2$sl,
    sig_tempera=(1-p_value_tempera)*100
    #sig2=1-2*(1-pnorm(abs((S-1)/(sqrt(54*53*(2*54+5)/18)))))
  )

perStation <- perStation %>%  
  mutate(
    slope_schneeh=summary(schneeh1)$coeff[2],
    intercept_schneeh=summary(schneeh1)$coeff[1],
    S_schneeh= schneeh2$S,
    varS_schneeh = schneeh2$varS,
    p_value_schneeh=schneeh2$sl,
    sig_schneeh=(1-p_value_schneeh)*100
  )
  
perStation <- perStation %>%  
  mutate(
    slope_schneet=summary(schneet1)$coeff[2],
    intercept_schneet=summary(schneet1)$coeff[1],
    S_schneet= schneet2$S,
    varS_schneet = schneet2$varS,
    p_value_schneet=schneet2$sl,
    sig_schneet=(1-p_value_schneet)*100
  ) 
```

See the results e.g. of the calculated significance values in percent:

```{r}
perStation %>%
  select(
    id, sig_tempera, sig_schneeh, sig_schneet
  )
```

Now, we prepare the data that is written to the output-files.

```{r}
stations <- stations %>%
  left_join(perStation, by = c("key" = "id")) %>%
  mutate(
    temperaS = sig_tempera,
    schneehS = sig_schneeh,
    schneetS = sig_schneet,
    trendEnde = substring(trendEnde, 1, 4)
  ) %>%
  select(key,name,hoehe,trendEnde,temperaS,schneehS,schneetS)
  
trends <- trends %>%
  left_join(perStation, by = c("id" = "id")) %>%
  mutate(
    tempera = as.numeric(substring(date, 1, 4)) * slope_tempera + intercept_tempera,
    schneeh = as.numeric(substring(date, 1, 4)) * slope_schneeh + intercept_schneeh,
    schneet = as.numeric(substring(date, 1, 4)) * slope_schneet + intercept_schneet
  ) %>%
  select(
    id, date, tempera, schneeh, schneet
  )

trends
```

Finally, we write the results to the CSV-files: 

```{r}
write.csv(trends, "./output/trends.csv", quote=FALSE, row.names=FALSE)
write.csv(stations, "./output/stationIDs.csv", quote=FALSE, row.names=FALSE)
file.copy("./input/klima.csv", "./output/klima.csv", overwrite=TRUE)
```